Kepler's Demographic Cliff: Ages of Young Stars and Planets in the Kepler Field

### 2024-05-02 paper/ms.tex
update ms

Our 
three main conclusions with respect to our original questions are as follows.
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
  %FIXME: 120 number
  \item From the rotation rates of Kepler planet-hosting stars, we
    have identified  \nplyounggyrotwosigma\ planets younger than
    1\,Gyr at 2$\sigma$, and \nplyounggyro\ planets with median ages
    below 1\,Gyr.
    This implies that there are now $\approx$120-170 known
    sub-gigayear planets with well-constrained ages.
  %
  \item The age distributions of both the Kepler
    target stars and the known Kepler planet-hosts shows a
    demographic cliff.  
    There are \ratioobtoybstars\ times as many ``old'' (2-3\,Gyr)
    stars in the Kepler field as ``young'' (0-1\,Gyr) stars.
    The star formation rate today is \ratiosfr$\pm$\uncratiosfr\ times
    lower than it was three billion years ago.
  %
  \item The implication of the latter age distribution is that rather
    than expecting $\approx$500 sub-gigayear exoplanets, we should
    instead expect $\approx$250.
    We have therefore gone from a factor of ten discrepancy to a
    factor of at most two.
%    The issue is therefore mostly, although perhaps not entirely, resolved.
%    Some of the remaining missing young
%    transiting planets may be in the Kepler data.  Many more are
%    likely in K2 and TESS.
\end{enumerate}

% 4 CepHer, 2 MELANGE3, 6 Theia520, 2 NGC6811.

@@

\facilities{
  Gaia \citep{Gaia_DR3_2022},
  Kepler \citep{Borucki10},
  Keck:I (HIRES).
%  TESS \citep{ricker_transiting_2015},
%  NGTS \citep{Wheatley_2018}
}

@@

\startlongtable
\begin{deluxetable*}{llllllllrrrrrrl}
  %\tabletypesize{\footnotesize}
  \tabletypesize{\scriptsize}
  \tablecaption{Ages of Kepler planets and planet candidates.  This
  version of the table is truncated to include only the youngest confirmed Kepler planets,
  sorted by the minimum of either the rotation or lithium-based
  age.  The
  full machine-readable table contains ages and age limits for
  \nnonfopkoissomeageinfo\ non-false positive KOIs with MES$>$10.  
  A shell script to decode the $Q_{\rm gyro}$ quality flag is {\bf \href{https://gist.github.com/lgbouma/20368253f1a98da1b39cf32fdda0be13}{available online}}.
  \label{tab:planets}}
  %\toprule
  %\midrule
  %\endhead
  \startdata
  KOI & Kepler &  $T_{\rm eff}$ & $P_{\rm rot}$ & EW$_{\rm Li}$ &
  $t_{\rm gyro}$ & $t_{\rm Li} $ & Consistent? & $R_{\rm p}$ & $P_{\rm orb}$ &
  $Q_{\rm planet}$ & $Q_{\rm gyro}$ &  Spec? & Comment \\
  -- &   -- & K & days & m\AA & Myr & Myr &  str &   Earths &    days &       int  & int & bool & -- \\
  \hline
  \input{table_subgyr_confirmed} %\\
% \rule
%  \input{table_subgyr_candidate}
  \enddata
  \tablecomments{``Confirmed'' planets appear in the machine-readable version before``candidate'' planets.
    The bit quality flags for the rotation-based ages,
    $Q_{\rm gyro}$, are described in Section~\ref{subsec:flags}.
    Concisely summarized, they are:
    Bit 0: $T_{\rm eff}/{\rm K} \in [ 3800-6200]$ ?
    Bit 1: $\log g$$<$4.2?
    Bit 2: $M_{\rm G}$<3.9 or $M_{\rm G}>8.5$?
    Bit 3: In KEBC?
    Bit 4: Large $d_{\rm xm,Kep-Gaia}$?
    Bit 5: Confused Kep-Gaia crossmatch?
    Bit 6: Gaia DR3 non-single star?
    Bit 7: RUWE$>$1.4?
    Bit 8: Crowded?
    Bit 9: Far from main-sequence?
    Bit 10: \citetalias{Santos_2021} CP/CB?
    Bit 11: $P_{\rm rot}$ not in the homogeneous
    \citetalias{Santos_2019} \citetalias{Santos_2021} sample?
    As an example, Kepler-1627 has $Q_{\rm gyro}$ flagged with bit 10 and bit 7.
    The analogous planet quality bit-flag, $Q_{\rm planet}$, has the
    following meaning.
    Bit 0: Candidate not reliable?
    Bit 1: MES$<$10?
    Bit 2: Grazing?
    For a star to have a high likelihood of being ``reliable for gyrochronology'' we suggest $Q_{\rm star}$ bits
    zero to nine to not be raised, and for a planet to be ``reliabe'', we suggested $Q_{\rm planet}$ to be zero.
  }
\end{deluxetable*}

>>>>

Our three main conclusions with respect to our original questions are
as follows.

\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
  %
  \item The age distributions of both the Kepler target stars and the
    known Kepler planet-hosts shows a demographic cliff.  There are
    \ratioobtoybstars\ times as many ``old'' (2-3\,Gyr) stars in the
    Kepler field as ``young'' (0-1\,Gyr) stars.  The star formation
    rate today is \ratiosfr$\pm$\uncratiosfr\ times lower than it was
    three billion years ago.
  %
  \item Rather than expecting $\approx$500 exoplanets younger than one
    billion years, the rotation-based age distribution of stars in the
    Kepler field implies that we should instead expect $\approx$250.
  %
  %FIXME: 120 number
  \item Using rotation-based ages, we have reported the ages of
    \nplyounggyrotwosigma\ Kepler planets younger than 1\,Gyr at 2$\sigma$,
    and \nplyounggyro\ planets with median ages below 1\,Gyr.
    Joining these planets against the existing literature yields
    $\approx$120-170 known sub-gigayear planets with well-constrained
    ages.  We have therefore gone from a factor of ten discrepancy to
    a factor of at most two.
\end{enumerate}

@@

\facilities{
  Gaia \citep{Gaia_DR3_2022},
  Kepler \citep{Borucki10},
  Keck:I (HIRES).
}

@@

\startlongtable
\begin{deluxetable*}{llllllllrrrrrrl}
  \tabletypesize{\scriptsize}
  \tablecaption{Ages of Kepler planets and planet candidates.  This
  version of the table is truncated to include only the youngest
  confirmed Kepler planets, sorted by the minimum of either the
  rotation or lithium-based age.  The full machine-readable table
  contains ages and age limits for \nnonfopkoissomeageinfo\ non-false
  positive KOIs with MES$>$10.  
  A shell script to decode the $Q_{\rm gyro}$ quality flag is {\bf \href{https://gist.github.com/lgbouma/20368253f1a98da1b39cf32fdda0be13}{available online}}.
  \label{tab:planets}}
  %\toprule
  %\midrule
  %\endhead
  \startdata
  KOI & Kepler &  $T_{\rm eff}$ & $P_{\rm rot}$ & EW$_{\rm Li}$ &
  $t_{\rm gyro}$ & $t_{\rm Li} $ & Consistent? & $R_{\rm p}$ & $P_{\rm orb}$ &
  $Q_{\rm planet}$ & $Q_{\rm gyro}$ &  Spec? & Comment \\
  -- &   -- & K & days & m\AA & Myr & Myr &  str &   Earths &    days &       int  & int & bool & -- \\
  \hline
  \input{table_subgyr_confirmed} %\\
% \rule
%  \input{table_subgyr_candidate}
  \enddata
  \tablecomments{``Confirmed'' planets appear in the machine-readable version before``candidate'' planets.
    The bit quality flags for the rotation-based ages,
    $Q_{\rm gyro}$, are described in Section~\ref{subsec:flags}.
    Concisely summarized, they are:
    Bit 0: $T_{\rm eff}/{\rm K} \in [ 3800-6200]$ ?
    Bit 1: $\log g$$<$4.2?
    Bit 2: $M_{\rm G}$<3.9 or $M_{\rm G}>8.5$?
    Bit 3: In KEBC?
    Bit 4: Large $d_{\rm xm,Kep-Gaia}$?
    Bit 5: Confused Kep-Gaia crossmatch?
    Bit 6: Gaia DR3 non-single star?
    Bit 7: RUWE$>$1.4?
    Bit 8: Crowded?
    Bit 9: Far from main-sequence?
    Bit 10: \citetalias{Santos_2021} CP/CB?
    Bit 11: $P_{\rm rot}$ not in the homogeneous
    \citetalias{Santos_2019} \citetalias{Santos_2021} sample?
    As an example, Kepler-1627 has $Q_{\rm gyro}$ flagged with bit 10 and bit 7.
    The analogous planet quality bit-flag, $Q_{\rm planet}$, has the
    following meaning.
    Bit 0: Candidate not reliable?
    Bit 1: MES$<$10?
    Bit 2: Grazing?
    For a star to have a high likelihood of being ``reliable for
    gyrochronology'' we suggest $Q_{\rm star}$ bits zero to nine to
    not be raised, and for a planet to be ``reliabe'', we suggested
    $Q_{\rm planet}$ to be zero.
  }
\end{deluxetable*}
